Software Architectures for Integrating Workflow Engines in Science Gateways

### 2016-05-13 paper.tex
done 3.3

\paragraph{Specific features} Meta-workflows are not supported by default
(\texttt{O$_1$}=1).  
\note{sil}{why not? if the engine is ws-pgrade it can. i really think that meta-workflows is a property exclusively of the engine. you can see that in fig 2.V interaction *6}
Debugging is not an issue since the science
gateway can retrieve any information from the workflow engine directly
(\texttt{O$_2$}=0).

@@

\paragraph{Extensibility} Adding a new type of workflow engine
requires to implement the corresponding workflow service, to modify
interaction \texttt{d}, and to implement interactions \texttt{b} and
\texttt{$c_2$} (\texttt{E$_1$}=4). New engine versions can be added by
updating the workflow service without modifying any interaction or
component (\texttt{E$_2$}=0). Updating the engine version in a
workflow service does not count in \texttt{E$_2$} since the only goal of this component
is to wrap the engine.  New workflows are added in the science gateway
or in the workflow engine through interactions \texttt{a$_1$} and
\texttt{a$_2$} (\texttt{E$_3$}=2). Adding a new type of infrastructure
requires updates in interactions \texttt{a$_1$}, \texttt{b},
\texttt{$c_1$} and \texttt{$c_2$} (\texttt{E$_4$}=4).

\paragraph{Scalability} The service architecture supports multiple
engine instances through multiple workflow services. In VIP for
instance, this feature has been available from release 1.17. A basic
load-balancing mechanism is available that sends new workflow
executions to the engine instance which has the least active
executions. To avoid ``black-hole'' syndromes created by failing
engine instances, engine instances are automatically disabled when
workflows cannot be submitted to them. Adding a new engine instance,
however, requires manual intervention to declare the new instance in
the science gateway (\texttt{S$_1$}=1). Consequently, elastic engines
are difficult to implement because they require a mechanism to update
the science gateway configuration when a new engine instance is
available. Distributing the execution of a single workflow in multiple
engines is usually not possible unless the workflow engine has
specific abilities (\texttt{S$_2$}=1). The scheduling of tasks on the
infrastructure is as complex as in any other architecture since the
workflow engine might implement any kind of scheduling policy
(\texttt{S$_3$}=0).

\paragraph{Specific features} Meta-workflows are not supported by default
(\texttt{O$_1$}=1). Fine-grained debugging information is usually easy
to obtain since the workflow service provides direct access to the
engine (\texttt{O$_2$}=0).

>>>>

\paragraph{Specific features} Meta-workflows are not supported by default
(\texttt{O$_1$}=1).  
\note{sil}{why not? if the engine is ws-pgrade it does. i really think that meta-workflows (and nested workflows) is a property exclusively of the engine. you can see that in fig 2.V interaction *6}
Debugging is not an issue since the science
gateway can retrieve any information from the workflow engine directly
(\texttt{O$_2$}=0).

@@

\paragraph{Extensibility} Adding a new type of workflow engine
requires implementing the corresponding workflow service,  modifying
interaction \texttt{d}, and  implementing interactions \texttt{b} and
\texttt{$c_2$} (\texttt{E$_1$}=4). New engine versions can be added by
updating the workflow service without modifying any interaction or
component (\texttt{E$_2$}=0). Updating the engine version in a
workflow service does not count in \texttt{E$_2$} since the only goal of this component
is to wrap the engine.  New workflows are added in the science gateway
or in the workflow engine through interactions \texttt{a$_1$} and
\texttt{a$_2$} (\texttt{E$_3$}=2). Adding a new type of infrastructure
requires updates in interactions \texttt{a$_1$}, \texttt{b},
\texttt{$c_1$} and \texttt{$c_2$} (\texttt{E$_4$}=4).

\paragraph{Scalability} The service architecture in principle supports multiple
engine instances through multiple workflow services. In VIP for
instance, this feature has been available from release 1.17. A basic
load-balancing mechanism is available that sends new workflow
executions to the engine instance that has the least active
executions. To avoid ``black-hole'' syndromes created by failing
 instances, engine instances are automatically disabled when
workflows cannot be submitted to them. Adding a new engine instance,
however, requires manual intervention to declare the new instance in
the science gateway (\texttt{S$_1$}=1). Consequently, elastic engines
are difficult to implement because they require a mechanism to update
the science gateway configuration when a new engine instance is
available. Distributing the execution of a single workflow in multiple
engines is usually not possible unless the workflow engine has
specific abilities (\texttt{S$_2$}=1). 
\note{sil}{is this score for VIP? i think that the score depends on the implementation of the science gateway, which could adopt some elastic mechanism within. so we should not say s_2=1 for the architecture}
The scheduling of tasks on the
infrastructure is as complex as in any other architecture since the
workflow engine might implement any kind of scheduling policy
(\texttt{S$_3$}=0).
\note{sil}{see comment in section 3.1}

\paragraph{Specific features} Meta-workflows are not supported by default
(\texttt{O$_1$}=1). 
\note{sil}{see comment in section 3.1}
Fine-grained debugging information is usually easy
to obtain since the workflow service provides direct access to the
engine (\texttt{O$_2$}=0).
\note{sil}{this is really not the case. the fact that the wf service is a separate component that hides information is a big problem for debugging at the science gateway level. 
we had a paper about this:
	@inproceedings{olabarriaga2014scientific,
		Author = {Olabarriaga, Silvia and Pierantoni, Gabrielle and Taffoni, Giuliano and Sciacca, Eva and Jaghoori, Mahdi and Korkhov, Vladimir and Castelli, Giuliano and Vuerli, Claudio and Becciani, Ugo and Carley, Eoin and others},
		Booktitle = {e-Science (e-Science), 2014 IEEE 10th International Conference on},
		Organization = {IEEE},
		Pages = {298--305},
		Title = {Scientific Workflow Management--For Whom?},
		Year = {2014}}
	}
